---
title: "Reproducible Research Project 1"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data

First we need to load csv file in data variable (Reader should set working directory before starting):
```{r}
setwd("F:/Online Courses/Reproducible Research/Porject 1/repdata_2Fdata%2Factivity")
data<-read.csv("activity.csv")

```


## Preprocessing of data

Change date format into date 
```{r,echo=TRUE}

data$date<-as.Date(data$date)

```
## Histogram without NA values

Histogram of total number of steps taken on each day is shown below:

```{r , echo=TRUE}

sum_steps<-aggregate(data$steps,by=list(data$date),FUN=sum,na.rm=TRUE) 

hist(sum_steps$x, 
      breaks=seq(from=0, to=25000, by=2500),
      col="yellow", 
      xlab="Total number of steps", 
      ylim=c(0, 20), 
      main="Histogram of the total number of steps taken each day\n(NA removed)")
 

```


## Mean and Median of Steps

Mean and median number of steps taken each day
```{r}

mean(sum_steps$x)


median(sum_steps$x)


```
## Time Series Plot

Time series plot of the average number of steps taken
```{r,echo=TRUE}

avg_steps<-aggregate(data$steps,by=list(data$interval),FUN=mean,na.rm=TRUE)

colnames(avg_steps)<-c("interval","steps")

library(ggplot2)

ggplot(aes(x=interval,y=steps),data=avg_steps)+geom_line()


```

## Maximum Average 5 minute interval

The 5-minute interval that, on average, contains the maximum number of steps
```{r}

avg_steps[avg_steps$steps==max(avg_steps$steps),1]


```

## Imputing NA

#### Here is code to describe and show a strategy for imputing missing data

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA's)
```{r}

sum(is.na(data$steps))


```


Replace NA values with the mean of the steps
```{r}

data$steps[is.na(data$steps)]<-mean(data$steps,na.rm=TRUE)


```
Here are some rows of new data set
```{r}

head(data)

```
## Histogram with Repaced NA values

Histogram of total number of steps taken on each day is shown below:

```{r , echo=TRUE}

sum_steps<-aggregate(data$steps,by=list(data$date),FUN=sum,na.rm=TRUE) 

hist(sum_steps$x, 
      breaks=seq(from=0, to=25000, by=2500),
      col="yellow", 
      xlab="Total number of steps", 
      ylim=c(0, 30), 
      main="Total number of steps taken each day\n(NA replaced by mean)")
 

```

Mean and median number of steps taken each day after replacing NA values with mean
```{r}

mean(sum_steps$x)


median(sum_steps$x)


```
##  Differences in activity patterns between weekdays and weekends

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
# Convert date into weekdays

data$days=tolower(weekdays(data$date))

#Now categorised days into weekend and weekdays

data$day_type<-ifelse(data$days=="saturday"|data$days=="sunday","weekend","weekday")

#Take mean steps taken on weekend or weekday in the intervals

avg_steps<-aggregate(data$steps,by=list(data$interval,data$day_type),FUN=mean,na.rm=TRUE)

colnames(avg_steps)<-c("interval","day_type","steps")

# Create panel plot between average steps and interval seperated by day type

ggplot(aes(x=interval,y=steps),data=avg_steps)+geom_line()+facet_wrap(~avg_steps$day_type)

```
